Label object representation and manipulation with ITK

### 2007-09-24 article/Article.tex
fix type usage in python example
darcs-hash:20070921105546-2fc9d-49ae828b652d2952bf9d53c8e591d099f202522e.gz
update article
darcs-hash:20070924095227-2fc9d-4bf53cd90d05a78694941f7e0231533beb6ceafa.gz

Richard Beare has recently introduced a new filter to efficiently labelize the
connected component with ITK, and has also proposed to use the run-length
encoding used in that filter to implement some of the most useful binary
mathematical morphology operators: the opening by attribute.
Following that idea, and after have searched a way to use the ITK's spatial
objects for this task, a new set of classes have been developed to represent and
manipulate the label images and the objects within them in ITK.
Those new classes have been used to implement several label images
manipulation based on object attributes, as well as the binary specialization of
some mathematical morphology filter already included in ITK, and not related to
the attribute of the objects. With those last filters, this contribution comes
with 49 new classes, and should greatly enhance the binary mathematical
morphology in ITK.

@@

\begin{itemize}
  \item {\em Minimum} is the minimum value in the feature image for the object. Its type is
the feature image pixel type.
  \item {\em MinimumIndex} is the index position in the image where the first minimum was
found. Its type is{\em Index< ImageDimension >}.
  \item {\em Maximum} is the maximum value in the feature image for the object. Its type is
the feature image pixel type.
  \item {\em MaximumIndex} is the index position in the image where the first maximum was
found. Its type is{\em Index< ImageDimension >}.
  \item {\em Mean} is the mean of the pixel values in the object. Its type is {\em double}.
  \item {\em Sum} is the sum of all the pixel values in the objects. Its type is {\em double}.
  \item {\em Sigma} is the standard deviation of the pixels values in the objects. Its type
is {\em double}.
  \item {\em Variance} is the variance of the pixels values in the objects. Its type is
{\em double}.
  \item {\em Median} is the median of the pixels values in the obejct. Its type is
{\em double}
  \item {\em CenterOfGravity} is the center of gravity of the object. It type is
{\em Point< double >}.
%   \item {\em CentralMoments}
%   \item {\em PrincipalMoments}
%   \item {\em PrincipalAxes}
  \item {\em Kurtosis} is the kurtosis of the pixel values in the objects. Its type is
{\em double}.
  \item {\em Skewness} is the skewness of the pixel values in the objects. Its type is
{\em double}.
  \item {\em PrincipalMoments} contains the principal moments. Its type is
{\em itk::Vector< double, ImageDimension >}.
  \item {\em PrincipalAxes} contains the principal axes of the object. Its type is
{\em itk::Matrix< double, ImageDimension, ImageDimension >}.
  \item {\em Elongation} is the elongation of the shape, computed as the ratio
of the largest principal moment by the smallest principal moment. Its value is greater
or equal to $1$· Its type si {\em double}.
\end{itemize}

@@

Finally, it has been chosen to develop a specific filter for the morphological
reconstruction. It would have been possible to implement the reconstruction
with the most common case (build the object collection, valuate the attributes
filter the object, and rebuild the image), but in order to make it compatible
with the {\em ShapeLabelObject} and {\em StatisticsLabelObject}, the reconstruction
filter filters the collection directly, without setting an attribute in the objects
\footnote{I'm not really pleased with that design though, and I'thinking about
reimplementing it using the generic attributes. It would have no impact on the
binary filters API.}.

@@

The general view of the previous section show a very common way to use those
classes. To make easier to use, some prebuilt classes have been made, to perform
the mini-pipeline:
\begin{itemize}
  \item creation of the {\em itk::LabelMap} from an {\em
itk::Image},
  \item valuation of the attribute(s) of the objects,
  \item filtering of the {\em itk::LabelMap},
  \item creation of an {\em itk::Image} from the filtered {\em
itk::LabelMap},
\end{itemize}
with a specific attribute.

@@

@@

@@

\section{Binary specialization of mathematical morphology filters}

@@

@@

Now, the spots:
\small \begin{verbatim}
spots = itk.ImageFileReader[ImageType].New(FileName="images/spots.png")
\end{verbatim} \normalsize
Mask the spot image to keep only the nucleus zone. The rest of the image is cropped, excepted a border of 2 pixels
\small \begin{verbatim}
maskSpots = itk.LabelMapMaskImageFilter.LI2IUC2.New(relabel, spots, Label=1,
    Crop=True, CropBorder=2)
\end{verbatim} \normalsize
A simple thresholding:
\small \begin{verbatim}
th = itk.BinaryThresholdImageFilter[ImageType, ImageType].New(maskSpots, LowerThreshold=110)
\end{verbatim} \normalsize
Now swith to the label collection representation, and compute the attribute values. This time, the input image is not a label image, but a binary one.
\small \begin{verbatim}
slabel = itk.BinaryImageToLabelMapFilter[ImageType, LabelMapType]
    .New(th)
sstats = itk.StatisticsLabelMapFilter[LabelMapType, ImageType]
    .New(slabel, nuclei)
\end{verbatim} \normalsize
we know there are for spots in the nubleus, so keep the 4 biggest spots. The other attribute are also usable - we may have chosen to keep the 4 brightest spots for example.
\small \begin{verbatim}
skeep = itk.ShapeKeepNObjectsLabelMapFilter[LabelMapType].New(
    sstats, Attribute='Size', NumberOfObjects=4)
\end{verbatim} \normalsize
Reoder the labels. The bigger objects first.
\small \begin{verbatim}
srelabel = itk.StatisticsRelabelLabelMapFilter[LabelMapType].New(
    skeep, Attribute='Size')
\end{verbatim} \normalsize
Finally, display the values we are interested in:
\begin{itemize}
  \item the nucleus number,
  \item the spot position,
  \item the mean value in the nucleus in the spot zone.
\end{itemize}

>>>>

Richard Beare has recently introduced a new filter to efficiently labelize the
connected component with ITK, and has also proposed to use the run-length
encoding used in that filter to implement some of the most useful binary
mathematical morphology operators: the opening by attribute.
Following that idea, and after have searched a way to use the ITK's spatial
objects for this task, a new set of classes have been developed to represent and
manipulate the label images and the objects within them in ITK.
Those new classes have been used to implement several label images
manipulation based on object attributes, as well as the binary and label specialization of
some mathematical morphology filters based on the morphological reconstruction. This contribution comes
with 57 new classes, and should greatly enhance the binary mathematical
morphology in ITK.

@@

\begin{itemize}
  \item {\em Minimum} is the minimum value in the feature image for the object. Its type is
the feature image pixel type.
  \item {\em MinimumIndex} is the index position in the image where the first minimum was
found. Its type is{\em Index< ImageDimension >}.
  \item {\em Maximum} is the maximum value in the feature image for the object. Its type is
the feature image pixel type.
  \item {\em MaximumIndex} is the index position in the image where the first maximum was
found. Its type is{\em Index< ImageDimension >}.
  \item {\em Mean} is the mean of the pixel values in the object. Its type is {\em double}.
  \item {\em Sum} is the sum of all the pixel values in the objects. Its type is {\em double}.
  \item {\em Sigma} is the standard deviation of the pixels values in the objects. Its type
is {\em double}.
  \item {\em Variance} is the variance of the pixels values in the objects. Its type is
{\em double}.
  \item {\em Median} is the median of the pixels values in the obejct. Its type is
{\em double}
  \item {\em CenterOfGravity} is the center of gravity of the object. It type is
{\em Point< double >}.
  \item {\em Kurtosis} is the kurtosis of the pixel values in the objects. Its type is
{\em double}.
  \item {\em Skewness} is the skewness of the pixel values in the objects. Its type is
{\em double}.
  \item {\em PrincipalMoments} contains the principal moments. Its type is
{\em itk::Vector< double, ImageDimension >}.
  \item {\em PrincipalAxes} contains the principal axes of the object. Its type is
{\em itk::Matrix< double, ImageDimension, ImageDimension >}.
  \item {\em Elongation} is the elongation of the shape, computed as the ratio
of the largest principal moment by the smallest principal moment. Its value is greater
or equal to $1$· Its type si {\em double}.
\end{itemize}

@@

@@

The general view of the previous section show a very common way to use those
classes. To make them easier to use, some prebuilt classes have been made, to perform
the mini-pipeline:
\begin{itemize}
  \item creation of the {\em itk::LabelMap} from an {\em
itk::Image},
  \item valuation of the attribute(s) of the objects,
  \item filtering of the {\em itk::LabelMap},
  \item creation of an {\em itk::Image} from the filtered {\em
itk::LabelMap},
\end{itemize}
with a specific attribute.

Also, when using the {\em itk::ShapeLabelObject} or the {\em itk::StatisticsLabelObject}
class, we usually want them to be valuated. Some classes are provided to perform the
mini-pipeline:
\begin{itemize}
  \item creation of the {\em itk::LabelMap} from an {\em
itk::Image},
  \item valuation of the attributes of the objects
\end{itemize}
with the {\em itk::ShapeLabelObject} or the {\em itk::StatisticsLabelObject}.

@@

The filters to produce valuated attributes:

\begin{itemize}
  \item {\em itk::BinaryImageToShapeLabelMapFilter}
  \item {\em itk::BinaryImageToStatisticsLabelMapFilter}
\end{itemize}

The filters to fully hide the usage of the label objects:

@@

The filters to produce valuated attributes:

\begin{itemize}
  \item {\em itk::LabelImageToShapeLabelMapFilter}
  \item {\em itk::LabelImageToStatisticsLabelMapFilter}
\end{itemize}

The filters to fully hide the usage of the label objects:

@@

\section{Binary and label specialization of mathematical morphology filters}

All the following filters are using a morphological reconstruction, implemented internally
as an attribute opening with the {\em itk::AttributeLabelObject} class. Some binary filters
are implemented:

@@

and some label filters, useful when the label objects are connected, or when we
want to avoid loosing the label by using a binary filter. Because the notion
of reconstruction by erosion is difficult with labels, only a few filters
are implemented:

\begin{itemize}
  \item {\em itk::LabelReconstructionByDilationImageFilter}
%  \item {\em itk::LabelOpeningByReconstructionImageFilter}
\end{itemize}

@@

Now, the spots:
\small \begin{verbatim}
spots = itk.ImageFileReader[ImageType].New(FileName="images/spots.png")
\end{verbatim} \normalsize
Mask the spot image to keep only the nucleus zone. The rest of the image is cropped, excepted a border of 2 pixels
\small \begin{verbatim}
maskSpots = itk.LabelMapMaskImageFilter[LabelMapType, ImageType].New(relabel, spots, Label=1,
    Crop=True, CropBorder=2)
\end{verbatim} \normalsize
A simple thresholding:
\small \begin{verbatim}
th = itk.BinaryThresholdImageFilter[ImageType, ImageType].New(maskSpots, LowerThreshold=110)
\end{verbatim} \normalsize
Now swith to the label collection representation, and compute the attribute values. This time, the input image is not a label image, but a binary one.
\small \begin{verbatim}
slabel = itk.BinaryImageToLabelMapFilter[ImageType, LabelMapType]
    .New(th)
sstats = itk.StatisticsLabelMapFilter[LabelMapType, ImageType]
    .New(slabel, nuclei)
\end{verbatim} \normalsize
we know there are for spots in the nubleus, so keep the 4 biggest spots. The other attribute are also usable - we may have chosen to keep the 4 brightest spots for example.
\small \begin{verbatim}
skeep = itk.ShapeKeepNObjectsLabelMapFilter[LabelMapType].New(
    sstats, Attribute='Size', NumberOfObjects=4)
\end{verbatim} \normalsize
Reoder the labels. The bigger objects first.
\small \begin{verbatim}
srelabel = itk.StatisticsRelabelLabelMapFilter[LabelMapType].New(
    skeep, Attribute='Size')
\end{verbatim} \normalsize
Finally, display the values we are interested in:
\begin{itemize}
  \item the nucleus number,
  \item the spot position,
  \item the mean value in the nucleus in the spot zone.
\end{itemize}
